---
title: "Iris_selecting_variables"
author: "Iris Chow"
date: "2024-12-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
merged_data5 <- read_csv("merged_data5.csv")
```

```{r}
# Count the number of NA values in each column
na_counts <- colSums(is.na(merged_data5))
print(na_counts)

```
```{r}

rf_model <- randomForest(Yield_Mg_ha ~ K_Sat, data = cleaned_data, importance = TRUE)

```


# Random Forest




```{r}
# Install and load the necessary packages
#install.packages("randomForest")
library(randomForest)

# Assume your data is loaded in a dataframe called merged_data5
# Ensure categorical variables are factors
merged_data5[] <- lapply(merged_data5, function(x) if(is.character(x)) as.factor(x) else x)

# Fit a Random Forest model
set.seed(123) # For reproducibility
# Remove rows with NA values
cleaned_data <- na.omit(merged_data5)

cleaned_data <- cleaned_data %>%
                rename(K_Sat = '%K Sat') %>%
                rename(Ca_Sat = '%Ca Sat') %>%
                rename(Mg_Sat = '%Mg Sat') %>%
                rename(Na_Sat = '%Na Sat') %>%
                rename(Sand = '% Sand') %>%
                rename(Silt = '% Silt') %>%
                rename(Clay = '% Clay') 

# Example for target encoding
#install.packages("vtreat")
library(vtreat)

# Create a treatment plan
# Create a treatment plan for regression
treatment_plan <- designTreatmentsN(
  cleaned_data, 
  varlist = names(cleaned_data)[sapply(cleaned_data, is.factor)], 
  outcomename = "Yield_Mg_ha"
)


# Apply the treatment plan to the dataset
cleaned_data_treated <- prepare(treatment_plan, cleaned_data)

# Fit the Random Forest model
rf_model <- randomForest(Yield_Mg_ha ~ ., data = cleaned_data_treated, importance = TRUE)

# Get variable importance
importance_scores <- importance(rf_model)
important_variables <- importance_scores[order(-importance_scores[, 1]), ]


```

                                                          %IncMSE IncNodePurity
Hybrid_catN                                           307.1313484  52677.832983
Hybrid_catD                                           141.6601420  22163.960624
Hybrid_catP                                           103.6950681  12167.881109
Experiment_catN                                        37.6421663   1684.703053
Env_catN                                               36.3803781  48069.702448
Date_Harvested_catN                                    36.2750372  35978.275322
Experiment_catD                                        33.3555034   1186.867725
Date_Harvested_catD                                    33.2709189   2129.995393
Experiment_catP                                        32.9949842   1232.623386
Date_Planted_catN                                      23.2637611  22674.775954
Date_Harvested_catP                                    22.3372907   1750.180009
Experiment_lev_x_G2F_2020_21_PHZ51                     21.4623280    458.215713
Farm_catD                                              21.1902112    824.789128
Date_Planted_catD                                      20.4008062    943.085041
Date_Planted_catP                                      20.1481164    858.576176
Env_catP                                               19.1799885   1302.045998
Env_catD                                               19.0658077   1639.729926
Experiment_lev_x_G2F_2022_23_LH244                     17.7069651    305.916345
Field_Location_catD                                    17.4768179    614.320938
Experiment_lev_x_G2F_2018_19_PHT69                     16.8989313    187.884516
City_catD                                              16.3233849    486.047695
Experiment_lev_x_G2F_2020_21_PHP02                     15.5979590    180.724374
Farm_catN                                              15.4621498   9583.750088
Experiment_lev_x_G2F_2020_21_PHK76                     15.2122008    162.827821
Farm_catP                                              14.9331073    692.722304
Field_Location_catP                                    14.8685260    544.570345
Experiment_lev_x_G2F_2018_19_LH195                     13.9144715    224.029317
Field_Location_catN                                    12.0277046   3784.756545
City_catN                                              11.7707362   5992.452254
Field_Location_lev_x_NYH3                              11.1817539     79.152003
Date_Harvested_lev_x_11_slash_28_slash_22              10.6952200     31.775210
Date_Harvested_lev_x_11_slash_7_slash_23               10.5198038     29.610476

```{r}
# Get the importance scores
importance_scores <- importance(rf_model)

# Convert to a data frame for easier manipulation
importance_df <- as.data.frame(importance_scores)

# Add variable names
importance_df$Variable <- rownames(importance_df)

# Sort by %IncMSE (or IncNodePurity, based on your preference)
sorted_importance <- importance_df[order(-importance_df$`%IncMSE`), ]

# Select the top 50 variables
top_50_vars <- sorted_importance$Variable[1:50]
print(top_50_vars)

```
```{r}
# Subset the cleaned_data_treated to include only the top 50 variables
top_50_data <- cleaned_data_treated[, c("Yield_Mg_ha", top_50_vars)]

# View the new dataset
head(top_50_data)

```


# LASSO

```{r}
# Install and load the necessary packages
#install.packages("glmnet")
library(glmnet)

# Prepare the data for glmnet
x <- model.matrix(your_response_variable ~ ., merged_data5)[, -1]
y <- merged_data5$your_response_variable

# Fit a LASSO model
lasso_model <- cv.glmnet(x, y, alpha = 1)

# Get the coefficients of the best model
lasso_coef <- coef(lasso_model, s = "lambda.min")
important_variables <- rownames(lasso_coef)[lasso_coef[, 1] != 0]
print(important_variables)

```

# Chi-square
```{r}
# Function to perform chi-square test and return p-values
chi_square_test <- function(data, response) {
  p_values <- sapply(names(data), function(x) {
    if (is.factor(data[[x]])) {
      chisq.test(table(data[[x]], data[[response]]))$p.value
    } else {
      NA
    }
  })
  return(p_values)
}

# Apply the chi-square test
chi_square_p_values <- chi_square_test(merged_data5, "your_response_variable")
important_variables <- names(chi_square_p_values)[chi_square_p_values < 0.05]
print(important_variables)

```

# Correlation Analysis for Numeric Variables

```{r}
# Calculate correlations
correlations <- sapply(names(merged_data5), function(x) {
  if (is.numeric(merged_data5[[x]])) {
    cor(merged_data5[[x]], merged_data5$your_response_variable, use = "complete.obs")
  } else {
    NA
  }
})

# Select important variables based on absolute correlation values
important_variables <- names(correlations)[abs(correlations) > 0.2] # Adjust threshold as needed
print(important_variables)

```

