---
title: "Iris_Organising_data"
output: html_document
date: "2024-12-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this notebook, I am trying to merge all the datasets into one from the data_cleaning folder made by Graham. From each small dataset, I try to condense the dataset. Here is how I do it.

In trait_data, I removed "Replicate" "Block""Plot""Range" "Pass" "Hybrid_orig_name" "Hybrid_Parent1""Hybrid_Parent2" .
Then I merge selected_trait_data and env_cov data by Env.
For the weather_season data, there are one row for each day for each Env, so I condensed the weather_season data group by Env, and calculate the average for other rows. Then, I merge the weather dataset with selected_trait_data and env_cov data.
For meta_data, I selected Env,Treatment,City,Farm and merge with other df.
I obtained lat, and long from envirotype_input. I group by Env and calculated the average.
For soil_data, I select only Env, `%K Sat`, `%Ca Sat`, `%Mg Sat`, `%Na Sat`, `% Sand`, `% Silt`, `% Clay` and merge them all together by Env

This is how I got to merged_data5 and obtain the csv. As the file size is too large to put in github. I have uploaded it to our google drive. https://drive.google.com/drive/folders/1pWHllpwmzqtKPwm218Wt96MxhtxkEvhT?usp=drive_link


I would need help on start to fit a very simple model like random forest. Please try it and let me know if there's further adjustments needed in the merged_data5

Another task is to utilize the genomic data, if catherine and graham could continue to work on that.

thank you




```{r}
setwd("~/UIUCDigitalHand")
library(readr)
library(tidyverse)
library(dplyr)
library(lme4)
library(mice)
library(caret)
library(randomForest)
library(glmnet)
```
```{r}
#env_cov <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/env_cov.csv")
meta_data <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/meta_data.csv")
soil_data <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/soil_data.csv")
trait_data <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/trait_data.csv")
weather_season <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/weather_season.csv")
weather_year <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/weather_year.csv")
hybrid_genetics_PCs <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/hybrid_genetics_PCs.csv")
hybrid_env_PCs <- read_csv("~/UIUCDigitalHand/data_cleaning/training_data/hybrid_env_PCs.csv")
#envirotype_input <- read_csv("data_cleaning/output_data/envirotype_input.csv")
```

# Get hybrid from trait

```{r}
library(dplyr)
# Select multiple columns by names 
selected_trait_data <- select(trait_data, "Env" ,"Year","Hybrid","Field_Location","Experiment","Yield_Mg_ha") 
# Print the selected columns 
print(selected_trait_data)
```
# join with PCs by hybrid
```{r}

# Rename the PC columns
new_names <- names(hybrid_env_PCs)
pc_indices <- grep("PC", new_names)
new_names[pc_indices] <- paste0("env_", new_names[pc_indices])
names(hybrid_env_PCs) <- new_names

# Print the new column names
print(names(hybrid_env_PCs))

```

```{r}
merge1 <- full_join(selected_trait_data, hybrid_env_PCs, by = "Hybrid")
head(merge1)
```

# Merge with another PC table

```{r}
# Rename the PC columns
new_names <- names(hybrid_genetics_PCs)
pc_indices <- grep("PC", new_names)
new_names[pc_indices] <- paste0("genetics_", new_names[pc_indices])
names(hybrid_genetics_PCs) <- new_names

# Print the new column names
print(names(hybrid_genetics_PCs))
```

```{r}
merge2 <- full_join(merge1, hybrid_genetics_PCs, by = "Hybrid")
head(merge2)
```

```{r}
# remove "LabID" and "Comments" in soil_data
# Load the dplyr package
library(dplyr)

# Assuming you have a data frame called soil_data
# Remove the columns "LabID" and "Comments"
soil_data <- soil_data %>% select(-LabID, -Comments)

# Print the updated data frame
print(soil_data)




```



```{r}
merge3<-full_join(merge2, soil_data, by = "Env")
```

```{r}
# Remove columns containing "Issue/comment"
meta_data <- meta_data[, !grepl("Issue/comment", names(meta_data))]

# Print the updated column names to verify
print(names(meta_data))

```
```{r}
# Assuming meta_data is your data frame

# Calculate average latitude and longitude
meta_data$Average_Latitude <- rowMeans(meta_data[, c(
  "Latitude_of_Field_Corner_#1 (lower left)",
  "Latitude_of_Field_Corner_#2 (lower right)",
  "Latitude_of_Field_Corner_#3 (upper right)",
  "Latitude_of_Field_Corner_#4 (upper left)"
)], na.rm = TRUE)

meta_data$Average_Longitude <- rowMeans(meta_data[, c(
  "Longitude_of_Field_Corner_#1 (lower left)",
  "Longitude_of_Field_Corner_#2 (lower right)",
  "Longitude_of_Field_Corner_#3 (upper right)",
  "Longitude_of_Field_Corner_#4 (upper left)"
)], na.rm = TRUE)

# Remove the corner columns
corner_columns <- c(
  "Latitude_of_Field_Corner_#1 (lower left)",
  "Latitude_of_Field_Corner_#2 (lower right)",
  "Latitude_of_Field_Corner_#3 (upper right)",
  "Latitude_of_Field_Corner_#4 (upper left)",
  "Longitude_of_Field_Corner_#1 (lower left)",
  "Longitude_of_Field_Corner_#2 (lower right)",
  "Longitude_of_Field_Corner_#3 (upper right)",
  "Longitude_of_Field_Corner_#4 (upper left)"
)

meta_data <- meta_data[, !names(meta_data) %in% corner_columns]

# Assuming meta_data is your data frame
# Remove columns 8, 9, 13, and 14
meta_data <- meta_data[, -c(8, 9, 13, 14)]
meta_data <- meta_data[, -c(8)]

# Print the updated column names to verify
print(names(meta_data))



```

```{r}
merge4 <- full_join(merge3, meta_data,by = "Env")
```

```{r}
names(weather_season)
```


# combine weather_season and weaher_year first

```{r}
# Load the dplyr package
library(dplyr)

# Combine the data frames by rows
combined_weather <- bind_rows(weather_season, weather_year)

# Print the combined data frame
print(combined_weather)

```
# simplify combined_weather

```{r}
# Load the dplyr package
library(dplyr)

# Remove the Date column
combined_weather <- combined_weather %>% select(-Date)

# Aggregate the data by Env and calculate the average for other columns
aggregated_weather <- combined_weather %>%
  group_by(Env) %>%
  summarise(across(everything(), mean, na.rm = TRUE))

# Print the aggregated data frame
print(aggregated_weather)

```



```{r}
trainingG2F <- full_join(merge4, aggregated_weather, by = "Env")
```

```{r}
write_csv(trainingG2F, "trainingG2F.csv")
```

# Testing

```{r}
testing_meta_data <- read_csv("~/UIUCDigitalHand/data_cleaning/testing_data/meta_data.csv")
testing_soil_data <- read_csv("~/UIUCDigitalHand/data_cleaning/testing_data/soil_data.csv")
testing_weather_season <- read_csv("~/UIUCDigitalHand/data_cleaning/testing_data/weather_season.csv")
testing_weather_year <- read_csv("~/UIUCDigitalHand/data_cleaning/testing_data/weather_year.csv")
template <- read_csv("~/UIUCDigitalHand/data_cleaning/testing_data/template.csv")
```

```{r}
library(dplyr)
# Select multiple columns by names 
submission_template <- select(template, "Env" , "Hybrid" ) 
# Print the selected columns 
print(submission_template)
```

```{r}
test1 <- left_join(submission_template, hybrid_env_PCs, by = "Hybrid")
```

```{r}
test2 <- left_join(test1, hybrid_genetics_PCs, by = "Hybrid")
```

```{r}
# Create column Field_Location from Env
# Assuming test2 is your data frame

# Create the Field_Location column by extracting the part of Env before the "_"
test2$Field_Location <- sub("_.*", "", test2$Env)

# Print the first few rows to verify
head(test2)

```

```{r}
# remove "LabID" and "Comments" in soil_data
# Load the dplyr package
library(dplyr)

# Assuming you have a data frame called soil_data
# Remove the columns "LabID" and "Comments"
testing_soil_data <- testing_soil_data %>% select(-LabID, -Comments)
testing_soil_data$Field_Location <- sub("_.*", "", testing_soil_data$Env)

# Print the updated data frame
print(testing_soil_data)
```

```{r}
test3<-left_join(test2, testing_soil_data, by = "Field_Location")
```

```{r}
testing_meta_data <- testing_meta_data[, !grepl("Issue/comment", names(testing_meta_data))]
# Calculate average latitude and longitude
testing_meta_data$Average_Latitude <- rowMeans(testing_meta_data[, c(
  "Latitude_of_Field_Corner_#1 (lower left)",
  "Latitude_of_Field_Corner_#2 (lower right)",
  "Latitude_of_Field_Corner_#3 (upper right)",
  "Latitude_of_Field_Corner_#4 (upper left)"
)], na.rm = TRUE)

testing_meta_data$Average_Longitude <- rowMeans(testing_meta_data[, c(
  "Longitude_of_Field_Corner_#1 (lower left)",
  "Longitude_of_Field_Corner_#2 (lower right)",
  "Longitude_of_Field_Corner_#3 (upper right)",
  "Longitude_of_Field_Corner_#4 (upper left)"
)], na.rm = TRUE)

# Remove the corner columns
corner_columns <- c(
  "Latitude_of_Field_Corner_#1 (lower left)",
  "Latitude_of_Field_Corner_#2 (lower right)",
  "Latitude_of_Field_Corner_#3 (upper right)",
  "Latitude_of_Field_Corner_#4 (upper left)",
  "Longitude_of_Field_Corner_#1 (lower left)",
  "Longitude_of_Field_Corner_#2 (lower right)",
  "Longitude_of_Field_Corner_#3 (upper right)",
  "Longitude_of_Field_Corner_#4 (upper left)"
)

testing_meta_data <- testing_meta_data[, !names(testing_meta_data) %in% corner_columns]

# Assuming testing_meta_data is your data frame
# Remove columns 8, 9, 13, and 14
testing_meta_data <- testing_meta_data[, -c(8, 9, 13, 14)]
testing_meta_data <- testing_meta_data[, -c(8)]

# Print the updated column names to verify
print(names(testing_meta_data))

testing_meta_data$Field_Location <- sub("_.*", "", testing_meta_data$Env)



```

```{r}
test4 <- left_join(test3, testing_meta_data,by = "Field_Location")
```

```{r}
# Load the dplyr package
library(dplyr)

# Combine the data frames by rows
testing_combined_weather <- bind_rows(testing_weather_season, testing_weather_year)
testing_combined_weather$Field_Location <- sub("_.*", "", testing_combined_weather$Env)

# Remove the Date column
testing_combined_weather <- testing_combined_weather %>% select(-Date)

# Aggregate the data by Env and calculate the average for other columns
testing_aggregated_weather <- testing_combined_weather %>%
  group_by(Field_Location) %>%
  summarise(across(everything(), mean, na.rm = TRUE))

testing_aggregated_weather <- testing_aggregated_weather %>% select(-Env)

# Print the combined data frame
print(testing_aggregated_weather)
```
```{r}
testingG2F <- left_join(test4, testing_aggregated_weather, by = "Field_Location")
```

```{r}
write_csv(testingG2F, "testingG2F.csv")
```

